Geometric variational problems

### 2020-11-13 NSF2020/Description/preliminarywork.tex
RJR
small edits
small edits to 4.3 and prelim work

\section{Preliminary work}
\subsection{Estimating two-dimensional elastic moduli}
The motivation for Specific Aims 1-3 stemmed from the successful demonstration of gradient driven
self-assembly of amphiphilic particles into two-dimensional micelles,
bilayer membranes, \cite{Fu2018_SIAM}, and recreation of the vesicles tank-treading phenomenon.

Our preliminary work on membrane elasticity 
is particulary encouraging because the effective bending, stretch, and tilt elastic moduli we calculated
were in excellent agreement with the experimental values (Table \ref{tab:moduli}).
The agreement is underscored by the fact we assigned reasonable physical values to 
the main model parameters of HAP energy, rather than tuning parameters.

\begin{table}[h!]
\begin{center}
  \begin{tabular}{|l|p{0.16\textwidth}|l|l|p{0.2\textwidth}|p{0.18\textwidth}}
    \hline
          & simulation type & HAP value        & experimental value & reference\\
    \hline \hline 
    $\KB$ & {\footnotesize uniform load with clamped condition} & $11 \pm 2.5$ \kBT & 10 \kBT                  & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & {\footnotesize bilayer stretching}  & $34$ \kBT $^{-2}$ & 30--40 \kBT $^{-2}$              & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & {\footnotesize tilt Dirichlet condition} & $12$ \kBT \; nm$^{-2}$ & 10 \kBT $^{-2}$  &  \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
\end{center}
\caption{\label{tab:moduli}Comparision of values elastic moduli from the experimental literature and values derived by HAP simulation.}
\end{table}

\subsection{Two-dimensional vesicle hydrodynamics in shear flow} 
%The motion of vesicles in shear flow is an important
%problem in the applied mathematics because simulations can reveal mechanical
%properties of membranes and lead to an enhanced understanding of 
%deformable particle laden flows \cite{Sinha15}. 
%
%
\begin{wrapfigure}[10]{r}{0.2\textwidth}
\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
\vspace{-8pt}
\caption{\label{fig:rupture} \footnotesize Rupture of tank-treading vesicle under strong shear flow.}
\end{wrapfigure}
%
To implement a vesicle in shear flow in the context of hydrophobic
potentials and mobility problem, we consider a shear flow in the
far-field $\mathbf{u}_{\infty} = Uy\mathbf{i}_x$ in the direction of the
$x$-axis (Figure \ref{fig:tanktreading}A). As illustrated in Figure
\ref{fig:tanktreading}, the particle based approach supports
inter-leaflet slip, and this can be used to determine inter-leaflet and
in-plane shear viscosities. 
%
%This field satisfies the linear Stokes system but does not give rise to a rigid motion at the particle interfaces. 
%To have a rigid motion, we change variables $\mathbf{u} = \tilde{\mathbf{u}}+ \mathbf{u}_{\infty}$ and 
%for the new field $\tilde{\mathbf{u}}$ vanishing at infinity we let 
%$\tilde{\mathbf{u}}|_{\partial P_i} = \mathbf{v}_i + \boldsymbol{\omega}_i \times (\mathbf{x} - \mathbf{a}_i)$ 
%where $(\mathbf{v}_i,\boldsymbol{\omega}_i)$ are the unknown translation and angular velocities of the 
%$i$th particle $P_i.$  
%
%\begin{wrapfigure}[17]{l}{0.4\textwidth}
%\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
%\caption{\label{fig:demixing} An initial assembly of small and 
%large particles spontaneous segregates into two smaller bodies. }
%\end{wrapfigure}
The HAP simulations show vesicle tank-treading. Under the external shear flow, the initially circular 
vesicle rotates in the clockwise direction. As the rate of rotation increases, the vesicle approaches
a steadily tank-treading ellipse. In Figure \ref{fig:tanktreading}B-D, the solid curves are ellipses fit to the particle centers
and midplane respectively. In the non-dimensionalized system, the particles have diameter 2, on the order of $\rho,$ 
and the vesicle diameter is about 14. 
%\todo[inline]{missing units. nm?}
Figure~\ref{fig:tanktreading}E shows the aspect ratio of the major to minor axes reaching an equilibrium value in the 
red and blue curves, yet oscillating in the high-shear rate (yellow) curve.
The tank-treading vesicle elongates and becomes more horizontal 
with an increase in flow rate or 
with a decrease in stiffness (effected by decreasing $\rho = 4$ to $\rho = 2$).

>>>>

\section{Preliminary work}
\label{sec:preliminary_work}
\subsection{Estimating two-dimensional elastic moduli} The motivation
for Specific Aims 1-3 stemmed from the successful demonstration of
gradient driven self-assembly of amphiphilic particles into
two-dimensional micelles, bilayer membranes~\cite{Fu2018_SIAM}, and
recreation of the vesicles tank-treading phenomenon.

Our preliminary work on membrane elasticity is particulary encouraging
because the effective bending, stretch, and tilt elastic moduli we
calculated were in excellent agreement with the experimental values
(Table~\ref{tab:moduli}).  The agreement is underscored by the fact we
assigned reasonable physical values to the main model parameters of HAP
energy, rather than tuning parameters.

\begin{table}[h!]
\begin{center}
  \begin{tabular}{|l|p{0.16\textwidth}|l|l|p{0.2\textwidth}|p{0.18\textwidth}}
    \hline
          & simulation type & HAP value        & experimental value & reference\\
    \hline \hline 
    $\KB$ & {\footnotesize uniform load with clamped condition} & $11 \pm 2.5$ \kBT & 10 \kBT                  & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & {\footnotesize bilayer stretching}  & $34$ \kBT $^{-2}$ & 30--40 \kBT $^{-2}$              & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & {\footnotesize tilt Dirichlet condition} & $12$ \kBT \; nm$^{-2}$ & 10 \kBT $^{-2}$  &  \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
\end{center}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{table}

\subsection{Two-dimensional vesicle hydrodynamics in shear flow} 
%The motion of vesicles in shear flow is an important problem in the
%applied mathematics because simulations can reveal mechanical
%properties of membranes and lead to an enhanced understanding of
%deformable particle laden flows \cite{Sinha15}. 
%
%
\begin{wrapfigure}[10]{r}{0.2\textwidth}
\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
\vspace{-8pt}
\caption{\label{fig:rupture} \footnotesize Rupture of tank-treading vesicle under strong shear flow.}
\end{wrapfigure}
%
To implement a vesicle in shear flow in the context of hydrophobic
potentials and mobility problem, we consider a shear flow in the
far-field $\mathbf{u}_{\infty} = Uy\mathbf{i}_x$ in the direction of the
$x$-axis (Figure \ref{fig:tanktreading}A). As illustrated in Figure
\ref{fig:tanktreading}, the particle based approach supports
inter-leaflet slip, and this can be used to determine inter-leaflet and
in-plane shear viscosities. 
%
%This field satisfies the linear Stokes system but does not give rise to a rigid motion at the particle interfaces. 
%To have a rigid motion, we change variables $\mathbf{u} = \tilde{\mathbf{u}}+ \mathbf{u}_{\infty}$ and 
%for the new field $\tilde{\mathbf{u}}$ vanishing at infinity we let 
%$\tilde{\mathbf{u}}|_{\partial P_i} = \mathbf{v}_i + \boldsymbol{\omega}_i \times (\mathbf{x} - \mathbf{a}_i)$ 
%where $(\mathbf{v}_i,\boldsymbol{\omega}_i)$ are the unknown translation and angular velocities of the 
%$i$th particle $P_i.$  
%
%\begin{wrapfigure}[17]{l}{0.4\textwidth}
%\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
%\caption{\label{fig:demixing} An initial assembly of small and 
%large particles spontaneous segregates into two smaller bodies. }
%\end{wrapfigure}
The HAP simulations show vesicle tank-treading. Under the external shear flow, the initially circular 
vesicle rotates in the clockwise direction. As the rate of rotation increases, the vesicle approaches
a steadily tank-treading ellipse. In Figure \ref{fig:tanktreading}B-D, the solid curves are ellipses fit to the particle centers
and midplane respectively. In the non-dimensionalized system, the particles have diameter 2, on the order of $\rho,$ 
and the vesicle diameter is about 14. 
%\todo[inline]{missing units. nm?}
Figure~\ref{fig:tanktreading}E shows the aspect ratio of the major to minor axes reaching an equilibrium value in the 
red and blue curves, yet oscillating in the high-shear rate (yellow) curve.
The tank-treading vesicle elongates and becomes more horizontal 
with an increase in flow rate or 
with a decrease in stiffness (effected by decreasing $\rho = 4$ to $\rho = 2$).

### 2020-11-13 NSF2020/Description/specificaim3_ver2.tex
small edits to 4.3 and prelim work

Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and this results in nearly-singular integrands. Quadrature
methods for these integrands has received a lot of attention in two and
three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
khayat_2005, schwab_1992, ying_2006, beale1, beale2, goodman_1990,
haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3,
helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. The trapezoid rule is the
workhorse for two-dimensional BIEs since it achieves spectral accuracy
when integrands are not nearly-singular~\cite{tre-wei2014}. To address
nearly-singular integrands of two-dimensional BIEs, we will use a {\em
barycentric quadrature rule} that only requires a slight modification of
the trapezoid rule~\cite{ioa-pap-per1991}. In its current form, this
method requires the layer potential to satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}, but the PIs will extend
the method to layer potentials for the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log|\mathbf{x} - \mathbf{y}|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x} - \mathbf{y}|$. Using this
decomposition, layer potentials involving $K_1(\mathbf{x},\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and layer potentials involving
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
$\mathbf{y}$. The proposed method is demonstrated in
Figure~\ref{fig:integrands} where the amplitude of the Fourier modes of
$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
plotted. The quadrature of the trapezoid rule is the sum of the Fourier
amplitudes not captured at the resolution.

@@

PI BQ developed a high-order adaptive time stepping method for
hydrodynamic suspensions~\cite{qua-bir2016} and it has served as a
robust method to simulate processes including mixing and adhesion in
suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}. The method uses
a single-step high-order time stepping method and a computationally
cheap estimate of the error. The proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order single-step method to achieve high-order accuracy.
To estimate the error, at each time step we will compute the total force
and torque of the system which is computationally cheap and physically
zero. The PIs' experience is that error estimates based on physical
constraints such as force- and torque-free conditions appropriately
adjust the time step size so that the dynamics are resolved without the
computational expense of techniques such as embedded Runge-Kutta
methods, step-doubling, and Richardson extrapolation.

@@

% ----------------------------------------------------------------------
\subsubsection{Fluctuating hydrodynamics}
\label{subsec:fluctuating}
Once algorithms that address quadrature, adaptive time stepping, and
contact are implemented, we will embark on incorporating fluctuating
hydrodynamics that are important even at scales of the some amphiphilic
particles. Bao et al.~\cite{Bao17,Bao18} developed two-dimensional BIE
formulations for Brownian rigid bodies. To satisfy the covariance of the
rigid body velocities, an ill-conditioned first-kind integral equation
method is used. The proposed work will consider a double-layer potential
formulation that has much better conditioning properties. This strategy
was attempted in~\cite{Bao18}, but they claim that a necessary
regularization of the covariance matrix leads to a drastic loss of
accuracy in numerical fluctuation-dissipation balance. However, they
only one regularization strategy was considered, and as PI BQ has
shown in previous work, the regularization choice significantly affects
the resulting physics~\cite{ong-chr-qua2017}. Therefore, different
regularization strategies will be investigated with the goal of
controlling the accuracy of the fluctuation-dissipation balance.

>>>>

Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and this results in nearly-singular integrands. Quadrature
methods for these integrands has received a lot of attention in two and
three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
khayat_2005, schwab_1992, ying_2006, beale1, beale2, goodman_1990,
haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3,
helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. The trapezoid rule is the
workhorse for two-dimensional BIEs since it achieves spectral accuracy
when integrands are not nearly-singular~\cite{tre-wei2014}. To address
nearly-singular integrands of two-dimensional BIEs, we will use a {\em
barycentric quadrature rule} that only requires a slight modification of
the trapezoid rule~\cite{ioa-pap-per1991}. In its current form, this
method requires the layer potential to satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}, but the PIs will extend
the method to layer potentials for the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log|\mathbf{x} - \mathbf{y}|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x} - \mathbf{y}|$. Using this
decomposition, layer potentials involving $K_1(\mathbf{x},\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and layer potentials involving
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
$\mathbf{y}$. The proposed method is demonstrated in
Figure~\ref{fig:integrands} where the amplitude of the Fourier modes of
$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
plotted. The quadrature error of the trapezoid rule is the sum of the
unresolved Fourier modes.

@@

PI BQ developed a high-order adaptive time stepping method for
hydrodynamic suspensions~\cite{qua-bir2016} and it has served as a
robust method to simulate processes including mixing and adhesion in
suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}. The method uses
a single-step high-order time stepping method and a computationally
cheap estimate of the error. The proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order single-step method to achieve high-order accuracy.
To estimate the error, at each time step we will compute the total force
and torque of the system which is computationally cheap and physically
zero. PI BQ's experience is that error estimates based on physical
constraints such as force- and torque-free conditions appropriately
adjust the time step size so that the dynamics are resolved without the
computational expense of techniques such as embedded Runge-Kutta
methods, step-doubling, and Richardson extrapolation.

@@

% ----------------------------------------------------------------------
\subsubsection{Fluctuating hydrodynamics}
\label{subsec:fluctuating}
Once algorithms that address quadrature, adaptive time stepping, and
contact are implemented, we will embark on incorporating fluctuating
hydrodynamics that are important even at scales of the some amphiphilic
particles. Bao et al.~\cite{Bao17,Bao18} developed two-dimensional BIE
formulations for Brownian rigid bodies. To satisfy the covariance of the
rigid body velocities, an ill-conditioned first-kind integral equation
method is used. The proposed work will consider a double-layer potential
formulation that has much better conditioning properties. This strategy
was attempted in~\cite{Bao18}, but they claim that a necessary
regularization of the covariance matrix leads to a drastic loss of
accuracy in numerical fluctuation-dissipation balance. However, only one
regularization strategy was considered, and as PI BQ has shown in
previous work, the regularization choice significantly affects the
resulting physics~\cite{ong-chr-qua2017}. Therefore, different
regularization strategies will be investigated with the goal of
controlling the accuracy of the fluctuation-dissipation balance.
